---
---crise 
title: "Speed and Max bout length estimation : tuto"
#subtitle: Part III - Tuberculosis Preventive Treatment Management
author: "Angelique Porciani"
date: last-modified
date-format: "[Last Updated on] DD MMMM, YYYY"
title-block-banner:  "#4E598C" 
format:
  html:
    code-fold: true
    html-math-method: katex
    code-tools: true
    self-contained: true
editor: visual
editor_options: 
  chunk_output_type: console
toc: true
execute:
  warning: false
  message: false
  cache: false
---

One of the parameters of interest is the duration of inactivity (or resting period). For estimate it, I need to compute the speed by timestep (mili or second) and apply a threshold to classify the larvae as moving or not. Here it is a detailled step of the analysis for individual tracked with rpi1 and for replicate 8. I apply all these step on all individual tracked in other script or at the end of this one.

The planned workflow is :

1.  Clean data (remove NA values), tracking loss in other script (NA_filter)
2.  Select individual keeped for analysis (Vid+LAM)
3.  Estimates of speed on 10fps time step and transform in cm/s
4.  Sampling by minutes
5.  Calculate activity variable (1 if \>0.3cm/s)
6.  Extract max bout length for activ and inactif state
7.  Save global df (all rpi and replicate, selected individual) in order to bind it with metadata table and analyse it in the script (rpt_viz_ana_indsel)

# 1. Load previously NA cleaned data

```{r}
library(MoveR)
library(readr)
library(tidyverse)
library(zoo)

# 1.Filter NA : other script, application of NA filter (loss of tracking) Import data already cleaned -----
# Pb with this approach : separation by tracklet do not conserve frame order... loss of timline information. 
rpi1_R08C<-data.table::fread(
  "./Data_larvae/data_larves_1hr/rpi1_10fps_R08_Corrected_cleanedData.csv.gz",
  sep = ";",
  dec = ".")

# selected individual for R08 
ID_selected <- read.csv("../ID_selected_MoveR.csv")
ID_rpi108 <- ID_selected %>% filter(str_detect(fullID3, ".108"))
selIDrpi108 <- as.character(ID_rpi108$fullID3)

# Select only individual that are alive until LAM 

rpi1_R08C_sub<- rpi1_R08C %>% filter(identity %in% selIDrpi108)

# Convert the data to a list of tracklets - here the data were saved as a varList object (a list of variable) a more simple structure than tracklets

trackDat_R08_sub_Cleaned <- MoveR::convert2Tracklets(rpi1_R08C_sub, by = "identity")
trackDat_R08 <- MoveR::convert2Tracklets(rpi1_R08C, by = "identity")
# set some useful information
trackDat_R08_sub_Cleaned <- MoveR::setInfo(trackDat_R08_sub_Cleaned, frameR = 10, scale = 1/37, imgRes = c(720, 1080)) 

```

Draw tracklets for the first 5 mins and selected individuals

```{r}

# draw tracklet 
MoveR::drawTracklets(trackDat_R08_sub_Cleaned,
                     imgRes = c(1080, 720),
                     #selId = selIDrpi108,
                     timeWin = list(c(0, 3000)), 
                     #colGrad = viridis::viridis(10))
                     colId = "identity", 
                     legend=F)

```

# 2. Speed estimation on 10 fps rate

And also mean speed (MA on 10 fps) and speed variance on 10 fps.

```{r}
trackDat_R08_sub_Cleaned <-
  MoveR::analyseTracklets(trackDat_R08_sub_Cleaned,
                          customFunc = list(
                            speed = function(x)
                              MoveR::speed(x,
                                           timeCol = "frame",
                                           scale = 1/37,# for in cm by second
                                           frameR=10, 
                                           timeU="s"), 
                            ### speed
                            slideMeanSpeed = function (y)
                              MoveR::slidWindow(y$speed,
                                                Tstep = 10, 
                                                statistic = "mean", 
                                                na.rm = T),
                            ### speed variance
                            slideVarSpeed = function (y)
                              MoveR::slidWindow(y$speed,
                                                Tstep = 10, 
                                                statistic = "var", 
                                                na.rm = T)
                          ))


```

Quick visualisation :

```{r}
plot(trackDat_R08_sub_Cleaned[[1]]$speed, type='l')
plot(trackDat_R08_sub_Cleaned[[1]]$slideMeanSpeed, type='l')

```

# 3. Resampling by 1 second

```{r}
# 2.B. Resampling tracklet to 1 second 
trackDat_R08_sub_Cleaned<-
  MoveR::analyseTracklets(trackDat_R08_sub_Cleaned,
                          customFunc = list(
                            # convert the time expressed in frame in second using a conversion factor of 25 frame per second
                            TimeSec = function(x)
                              x[["frame"]] / 10
                          ))

# resample the tracklets every 1 seconds
trackDat_R08_sub_Cleaned_byS <-
  MoveR::resampTracklets(trackDat_R08_sub_Cleaned, timeCol = "frame",  Tstep = 10)
```

Check size of time before and after sampling :

```{r}


## check the size of the tracklets
trackSize <- unlist(lapply(trackDat_R08_sub_Cleaned, function(x)
  nrow(x)))


## check the size of the tracklets after resampling
trackSize1s <- unlist(lapply(trackDat_R08_sub_Cleaned_byS, function(x)
  nrow(x)))

## Compare the tracklets size
cbind(trackSize, trackSize1s)
```

Draw trajectory by second

```{r}
MoveR::drawTracklets(trackDat_R08_sub_Cleaned_byS,
                     timeCol = "TimeSec",
                     imgRes = c(1080, 720),
                     #selId = selIDrpi108,
                     timeWin = list(c(0, 300)), 
                     #colGrad = viridis::viridis(10))
                     colId = "identity", 
                     legend=F)
```

Estimate activity

```{r}

trackDat_R08_sub_Cleaned_byS<-
  MoveR::analyseTracklets(trackDat_R08_sub_Cleaned_byS,
                          customFunc = list(
                            activity1 = function(x)
                            MoveR::activity1(x, 
                                             speedCol = "speed", 
                                             minSpeed = 0.3), 
                            MaxSpeed=function(x) max(x[["speed"]], na.rm=TRUE)
                          ))

# seuil de 0.3 cm/s as in animalTA for now and corresponding to mean larval length

```

# 4. Extract max bout length for activ and inactif state

Traitement des NA =\> replacement par la précedente valeur (discutable mais c'est ce qu'il font dans ethoscopy) a garder pour les HMM surtout pour les TAngle

```{r}
#function of bout extraction 
split_by_repetition <- function(vec) {
  if (sum(is.na(vec))>0){
    vec <- na.locf(vec)
  }
  if (sum(is.na(vec))==0){
    vec <- vec}
  # Identify the indices where consecutive numbers are broken or a repetition changes
  breaks <- c(0, which(diff(vec) != 0), length(vec))
  
  # Use those break points to split the vector into sublists
  result <- lapply(seq_along(breaks)[-length(breaks)], function(i) {
    vec[(breaks[i] + 1):breaks[i + 1]]
  })
  
  # now keep only ones corresponding to 1 
  filtered_active <- Filter(function(x) all(x == 1), result)
  
  # same for 0 
  filtered_inactive <- Filter(function(x) all(x == 0), result)
  
  res_Lb1<-max(unlist(lapply(filtered_active, length)))
  res_Lb_in<-max(unlist(lapply(filtered_inactive, length)))
  return(c(res_Lb1, res_Lb_in))
}

# apply 
res_Lb<-list()
df_res <- data.frame(Max_bout_activ=NA, Max_bout_inactiv=NA, fullID3=NA)

for (i in 1:length(trackDat_R08_sub_Cleaned_byS)){
  res_Lb[[i]]<-split_by_repetition(trackDat_R08_sub_Cleaned_byS[[i]]$activity1)
  df_res[i,] <- data.frame(Max_bout_activ=res_Lb[[i]][1], Max_bout_inactiv=res_Lb[[i]][2], fullID3=names(trackDat_R08_sub_Cleaned_byS)[i])
}

library(kableExtra)
df_res %>% kbl(caption = "Estimation of max bout length for active and inactive for individuals of rpi1 R08", 
               booktabs = T)%>%
	kable_styling(full_width = F)

str(df_res)
```
